\begin{thebibliography}{99}

\bibitem{newman2010}
M. E. J. Newman, \emph{Networks: An Introduction}. Oxford University Press, 2010.

\bibitem{easley2010}
D. Easley and J. Kleinberg, \emph{Networks, Crowds, and Markets: Reasoning About a Highly Connected World}. Cambridge University Press, 2010.

\bibitem{milgram1967}
S. Milgram, ``The small world problem,'' \emph{Psychology Today}, vol. 2, no. 1, pp. 60--67, 1967.

\bibitem{watts1998}
D. J. Watts and S. H. Strogatz, ``Collective dynamics of `small-world' networks,'' \emph{Nature}, vol. 393, pp. 440--442, 1998.

\bibitem{blondel2008}
V. D. Blondel, J.-L. Guillaume, R. Lambiotte, and E. Lefebvre, ``Fast unfolding of communities in large networks,'' \emph{Journal of Statistical Mechanics: Theory and Experiment}, vol. 2008, no. 10, P10008, 2008.

\bibitem{liben2007}
D. Liben-Nowell and J. Kleinberg, ``The link-prediction problem for social networks,'' \emph{Journal of the American Society for Information Science and Technology (JASIST)}, vol. 58, no. 7, pp. 1019--1031, 2007.

\bibitem{adamic2003}
L. A. Adamic and E. Adar, ``Friends and neighbors on the Web,'' \emph{Social Networks}, vol. 25, no. 3, pp. 211--230, 2003.

\bibitem{takac2012}
L. Takac and M. Zabovsky, ``Data analysis in public social networks,'' in \emph{International Scientific Conference and International Workshop Present Day Trends of Innovations}, 2012.

\bibitem{snap2014}
J. Leskovec and A. Krevl, ``SNAP Datasets: Stanford Large Network Dataset Collection,'' 2014. [Online]. Available: \url{https://snap.stanford.edu/data/}

\bibitem{networkx2008}
A. A. Hagberg, D. A. Schult, and P. J. Swart, ``Exploring network structure, dynamics, and function using NetworkX,'' in \emph{Proceedings of the 7th Python in Science Conference (SciPy 2008)}, Pasadena, CA, USA, 2008.

\bibitem{sklearn2011}
F. Pedregosa \emph{et al.}, ``Scikit-learn: Machine Learning in Python,'' \emph{Journal of Machine Learning Research}, vol. 12, pp. 2825--2830, 2011.

\bibitem{plotly}
Plotly Technologies Inc., ``Plotly Python Open Source Graphing Library.'' [Online]. Available: \url{https://plotly.com/python/}

\bibitem{streamlit}
Streamlit Inc., ``Streamlit Documentation.'' [Online]. Available: \url{https://docs.streamlit.io/}

\bibitem{backstrom2012}
L. Backstrom, P. Boldi, M. Rosa, J. Ugander, and S. Vigna, ``Four degrees of separation,'' in \emph{Proceedings of the 4th Annual ACM Web Science Conference (WebSci '12)}, pp. 33--42, 2012.

\bibitem{kwak2010}
H. Kwak, C. Lee, H. Park, and S. Moon, ``What is Twitter, a social network or a news media?'' in \emph{Proceedings of the 19th International Conference on World Wide Web (WWW '10)}, pp. 591--600, 2010.

\bibitem{albert2000}
R. Albert, H. Jeong, and A.-L. Barab\'{a}si, ``Error and attack tolerance of complex networks,'' \emph{Nature}, vol. 406, pp. 378--382, 2000.

\bibitem{kleinberg2000}
J. Kleinberg, ``The small-world phenomenon: An algorithmic perspective,'' in \emph{Proceedings of the 32nd Annual ACM Symposium on Theory of Computing (STOC '00)}, pp. 163--170, 2000.

\bibitem{linkedin2016}
LinkedIn Engineering, ``Measuring the LinkedIn professional network,'' LinkedIn Engineering Blog, 2016. [Online]. Available: \url{https://engineering.linkedin.com/}

\bibitem{barabasi1999}
A.-L. Barab\'{a}si and R. Albert, ``Emergence of scaling in random networks,'' \emph{Science}, vol. 286, no. 5439, pp. 509--512, 1999.

\bibitem{brandes2001}
U. Brandes, ``A faster algorithm for betweenness centrality,'' \emph{Journal of Mathematical Sociology}, vol. 25, no. 2, pp. 163--177, 2001.

\bibitem{fortunato2010}
S. Fortunato, ``Community detection in graphs,'' \emph{Physics Reports}, vol. 486, no. 3--5, pp. 75--174, 2010.

\bibitem{travers1969}
J. Travers and S. Milgram, ``An experimental study of the small world problem,'' \emph{Sociometry}, vol. 32, no. 4, pp. 425--443, 1969.

\bibitem{dodds2003}
P. S. Dodds, R. Muhamad, and D. J. Watts, ``An experimental study of search in global social networks,'' \emph{Science}, vol. 301, no. 5634, pp. 827--829, 2003.

\end{thebibliography}
